---
Title: Project One
output: html_document
---
# Data Project One"
## 100 points"
## DUE Sept 15 at 11:59 PM"

Below are the questions for the first data practical assignment. This project uses the FossilAnts.csv file" located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

(5 pts) Create a directory called `projects` and in it a subdirectory called `project_one`. Use `download.files` to get [this file](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd)"  and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv) and save them to the `project_one` directory."


#1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly."

```{r}
#Enter Your Answer Here"
library(tidyverse)
project_dat <- fossil_ants
glimpse(project_dat)

```

#2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer 
#(this is an opinion question - answers may vary)."

```{r}
# Answer here
No.It should be treated like a string of text similar to a name.  

```

#3.(5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character()."

```{r}
#Answer Here"
project_dat$reference_no<-as.character(project_dat$reference_no)
str(project_dat)

```

#4. (5 pts) Look at your data. What are the missing data values? In particular have a look at the `Tribe` column. In your opinion are these intelligent missing values for the dataset? Why or why not? If not how would you like to change them?

```{r}
#Answer here

Missing data values are basically cells within the dataset that dont have any data entered either due to error, lack of data point from the study usually. I dont think they are intelligent. Most of the column is NONE anyway so if there was supposed to be something, it would be less likely  that someone just forgot to input it because those are so rare in this column to begin with. 

I would set the empty cells to be equal to NONE like the rest. Or if i wanted to be safe, I may choose to set them equal to NA then remove them.
```

#5.(5 points) In the surveys dataset we have genus and species split between two columns. Here we have them combined. What are the pros and cons of the way we have recorded taxa in these two data sets."

```{r}
# Answer here"
We learned in class that its best to only have one type of information in each column. This is ideal so there is never any confusion about what the value of the cell is. When making graphs you have to specify a specific column name, and if you are working with a combined column, its so difficult to separate both genus and species in that example. 

```

#6.(5pts) Please look at the help page for the `separate` function."

```{r}
# Show how you would pull up the help"
?separate
```


#7.(10 pts) How could you separate one column into two? 

```{r}
#Answer here
separate(project_dat, col=taxon, into=c('genus', 'species'), sep='_')
```

Next we will test a hypothesis. Your hypothesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval. 75-100  30-0

#8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 

```{r}
#Answer Here"
I decided to make the column into a vector where I first counted the number of values that were between 75-100. Then I repeated that but counted the number of values that were between 30-0. It is a "simple" way to do it. I wanted to make some sort of pipe that would count both ranges at the same time but... I haven’t figured that out yet. 
```

#9. (15 pts) Perform the operations you described in (8). 

```{r}
sum(min_ma_df>= 0 & min_ma_df <= 30)

sum(min_ma_df>= 75 & min_ma_df <= 100)

#AMW: Elegant. I think you used less code than anyone else and got the same solution as everyone else.
```

#10. (5 pts) Do the results of your code support the hypothesis? 

```{r}
#Answer here
Nope. Our hypothesis state that there would be more specimens between 75 and 100 mya, however we now see that there are significantly more specimen present within the last 30 my. There were only 51 species that were dated between 75-100 mya compared to over 350 specimen s from 30 mya to present day. With this in mind, our hypothesis was not supported. combination of the three.

#11. (10 pts) Save the dataframe with the split taxon columns into a new directory called `project_one_data_output`. Save it as a csv file called column_separated.csv "

```{r}
# Answer here
column_separated.csv <- project_one_data_split
write.csv(project_one_data_split, "columns_separated.csv")
#Every time I try to alter the path, i get some many errors. I've tried many different ways but.. so far the only successful way is to save it to the working directory.. 
```

#12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help."

```{r}
#Answer here"
ggplot(column_separated.csv, aes(x = min_ma)) +
     geom_histogram(binwidth = 5)

```

#13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as large_font.pdf."

```{r}
large_font <- ggplot(column_separated.csv, aes(x = min_ma)) +
     geom_histogram(binwidth = 5) +
  theme(axis.text.x = element_text(size = 40)) +
  theme(axis.title = element_text(size = 40)) +
   theme(axis.text.y = element_text(size = 40))

ggsave("large_font.pdf", plot = large_font, width = 9, height = 7)

#it is saved in my project_one_data_output
```

#14. Finally produce a histogram of counts for each subfamily."

```{r}
ggplot(column_separated.csv, aes(x = subfamily)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#AmW: I like the angle.
```

# Graduate Students
1. You will do the above steps with the classroom dataset.

2. Next you will choose three data steps from above to perform with your own data. This could be grouping mutating cleaning NA values visualization or any combination of the three.

#I was not able to upload my own data from my laptop, so I just borrowed the butterfly data set. 

```{r}
butterfly_split <- separate(Butterfly_data, col=ButterflySpecies, into=c('genus', 'species'), sep=' ') 
na.omit(butterfly_split)
  glimpse (butterfly_split)

```

```{r}
ggplot(Butterfly_data, aes(x = ButterflySpecies)) +
geom_bar() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
Butterfly_data$Year<-as.character(Butterfly_data$Year)
str(Butterfly_data)

ggplot(Butterfly_data, aes(x = Year, fill = ButterflySpecies)) +
geom_bar() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(Butterfly_data, aes(x = Year, fill = ButterflySpecies)) +
  geom_bar(position = "dodge")
```

3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

```{r}
#Answer Here
I chose these combination of steps because I will most certainly need to know how to use these functions to organize my data frame for my research. In doing this, I realized that for ggplot, bar function, the x-axis doesn’t like to be a integer. So I was able to change it to a character and successfully obtained my goal in showing the data by year in my graph. Because timing is so important to my study, I will need to sort the data by date of collection. Overall, I wanted more practice making these graphs.

I anticipate making many bar plots for my thesis project, and the function they serve will be to show the change in hormone concentration over the course of the nesting season. Learning and practicing how to change what i am trying to visualize is vital.
```
