---
title: "Homework One"
output: html_document
---

## Homework One

### Due Aug. 25th at 5pm. Simply write your answers in and save in the homeworks folder in your Rstudio.

This is going to be a low-stakes effort to get you comfortable with RStudio and RMarkdown. First, you're going to download the RMarkdown Document for this homework. Copy and paste this command into the console of your RStudio instance.

```{r}
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/homework_1.Rmd", destfile = "/cloud/project/homeworks/homework_1.Rmd")
```

Can you find where the file downloaded to? Look at the help entry for `download.file` if you aren't sure.

Open the file and answer these two questions: 

1. (1pt) What does `download.file` do? 
```
Enter your answer here!
This function can be used to download a file from the Internet.
```

2. (1pt) Where did your HW1.Rmd file download to? 

```
Enter your answer here!
It was downloaded in the cloud in the projects folder and more specifically in the homeworks folder
```

Next, we're going to try to execute some code in an RMarkdown document. Press the green triangle.

```{r}
animals <- c("cat", "rat", "dog")
weights <- c(15, 1, 120)
animals[c(1,3)]
```

What just happened? Annotate the above code with two comments (each worth one point) explaining the code.

-When I pressed the green triangle, it printed the code. The last line of code was for a combination vector that is tasked with printing a list of a specified combination. 
-Additionally, it only printed "cat" and "dog", this is because on line 38, a combine vector was used to list only the first and third item in the initial animal list. ie. "Cat" and "Dog"

We can also embed plots in RMarkdown. The below code should embed a barplot. Not very useful, but let me know if you don't see one show up in your file.

```{r}
barplot(names.arg=animals, height=weights)
```

Did the plot show up OK?

```
Answer yes or no here
Yes.
```

## Getting to know you all

For the remainder of the assignment, I need to know three things.

1. What type of data will you be working with for your thesis or research project? For example: matrices of phylogenetic characters, spreadsheets of ecological observations, shapefiles from 3D scanning, etc. If you aren't actively doing research, what kinds of scientific questions do you find interesting?

I will be performing assays on all my plasma (Testosterone, E2, and VTG) and those results will be the bulk of my data. How ever I will have other biological data which includes both numerical data such as carapace length and categorical data such as the activity the turtle was in when we encountered her on the beach. 

#AMW: Cool.surveys<- read_csv ("data/portal_data_joined.csv")


2. What are you hoping to get out of class?

I am hoping to be competent in R and develop the ability to continue too use this tool long after leaving SELU. It's a very powerful tool and I want to be able to utilize it whenever that need arises for a multitude of uses. 

3. Are there topics that aren't on the schedule that you'd be interested to see? 

I am especially interested in data visualizations. I hope to learn how to create really beautiful ways to showcase a variety of data. 
