------------------------------------------------------------------------

title: "Homework Three" author: "April Wright" output: html_document

# Plotting and choices

## 10 points, due Sept. 8 at 5 PM

```{r}
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/homework_3.Rmd", destfile = "/cloud/project/homeworks/homework_3.Rmd")
```

0.  Read in the surveys.csv dataset. Remove NAs from the datasets.

1.  In this homework, we're going to explore how different ways of visualizing the same data. First, take a look at `geom_col()`. Group the surveys data by the species_id column and count how many of each species there are. Use this count data to make a bar plot of the counts per species.

```{r}
#Answer here

 ggplot(count_species, aes(x = species_id, y = count)) +
     geom_col() +
     labs(title = "Bar Plot of Species ID Counts", x = "Species_id", y = "Count")
```


2.  Take a look at the argument `fct_reorder`. It reorders variables on one or both axes. Try using this to plot the species in order from most to least members.

```{r}

 ggplot(count_species, aes(x = fct_reorder(species_id, -count), y = count)) +
     geom_bar(stat = "identity") +
     labs(title = "Bar Plot with Species ID counts in Descending Order of Value", x = "Species ID", y = "Count")
 
# Answer here
```

3.  What we canonically think of as a bar plot can also be made in R. These can have some interesting properties, such as being able to fill in bars by other aesthetics. Using the surveys data set, try to plot the number of members of each species, with the bar filled in by sex.

```{r, eval=FALSE}
ggplot(count_species_sex, aes(x = species_id, y = count, fill = sex)) +
  geom_col()


```

4.  `geom_col` accepts various arguments. Try `postion="dodge"` or position="stack". How does this change the plot and how you interpret it?

```{r}
ggplot(count_species_sex, aes(x = species_id, y = count, fill = sex)) +
  geom_col(position = "dodge")

ggplot(count_species_sex, aes(x = species_id, y = count, fill = sex)) +
  geom_col(position = "stack")

#Answer here.
#The stack graph shows more of a total view with both sexes stack on one another. Contrastingly, the dodge show them side by side and you can more clearly see how many of each sex there are. If you wanted to get a brief sense of the ratio of the sexes, i would maybe look at stacked. Whereas if you wanted to have a more clear representation of the actual numbers of each, I would look at dodge. 
```

5.  Even a simple histogram can lead to different interpretations. Make a histogram of hindfoot_lengths in the surveys dataset. Try choosing different binwidths. How does a high bindwith (like 100) this change your interpretation of how the data are distributed?

```{r}

ggplot(surveys, aes(x = hindfoot_length)) +
  geom_histogram(binwidth = 5)

ggplot(surveys, aes(x = hindfoot_length)) +
  geom_histogram(binwidth = 100)

ggplot(surveys, aes(x = hindfoot_length)) +
  geom_histogram(binwidth = 50)
#Answer here.
#I dont really understand why the count is changing so drastically. But basically, the binwidth refers to how wide the range is that you are looking at. The smaller the binwidth, the more detailed and finer the details in the histogram. I really dont like the look of the larger binwidths. I dont think I'd use that very often. It seems to distort the data alot. 
```

6.  Next, let's take a look at density plots. First, look at the help for `geom_density`. Density plots are like a smoothed histogram, mostly used for continuous data. But how density is calculated is done using what is called a kernel. To get a sense for what this means, try different kernel types. Some common ones are "triangular", "rectangular", and "gaussian." Try them out, and put the one you think best represents the data in the answer below.

```{r, eval=FALSE}

ggplot(surveys, aes(x = hindfoot_length)) +
  geom_density(fill = "blue", binwidth = 10, kernel ="rectangular")

#ggplot(surveys, aes(x = hindfoot_length)) +
 # geom_density(fill = "blue", binwidth = 10, kernel ="triangular")

#ggplot(surveys, aes(x = hindfoot_length)) +
 # geom_density(fill = "blue", binwidth = 10, kernel ="gaussian")

#I think the rectangular one best represent the data because it just shows more of the detail. For data I feel like the more detailed the graphic, the accurate the information that's being communicated. 
```

## Graduate Students:

A common set of journal figure requirements several of you submitted include the following:

-   PDF or PNG
-   At least 300 DPI
-   One column (80 mm) or two column (160mm) wide.

For each of your plots on the above homework, save figures meeting all requirements in your lastname_directory in a directory called `output.`

#I really struggled with this section. The saving the pdf to a new directory really tripped me up and I kept getting error messgaes. I did have some success saving one pdf but not the way I think you wanted it. I will come by your office this upcoming week to figure it out and discuss it with you. Thank you!

#AMW: Sounds good!

As a longer-term goal, you each picked a few figures. See if you can find the data that the authors used to make the figures. One of the best ways to learn to make good figures in your field is to copy from what's already published.

What data did the author's use?
```         
Answer here
Reproductive biomarker concentrations from assays (T, E2, and VTG), weight in kg, time such as month of week in the nesting season.
```

Is the data you're using for class similar to these data?
```         
Answer here
It is similar in that we work with a lot of quantitative data and also categories. For my data, I anticipte grouping my data by time (month), maybe size as well.  
```

Could you process your data to be similar to these data?
```         
Answer here
I think so. What we are doing seems like it will translate very well to my own project since I am simply making conections between hormone concentrations and weight and time.
```

Finally, in the final project in this class, you'll need to produce an R package with five functions:

-   One or two for data cleaning and/or data manipulation
-   One or two for a statistical test (ANOVA, linear models, other specialized analyses)
-   One or two for plotting
-   One of your choice for funsies.

In next week's homework, you'll be expected to pick and write one. So start thinking about one now.
